---
output: 
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# A Repository of TMB Training Materials

## TMB first steps
See [Getting Started with TMB](https://nsawtraining.github.io/TMBtraining/articles/00_01_Getting_Started.html) for software requirements and TMB package installation instructions. The [TMB Developer Resources](https://nsawtraining.github.io/TMBtraining/articles/00_00_TMB_Developer_Resources.html) provide a collection of useful resources from the TMB developers. 

<!-- badges: start -->

[![GitHub release (latest by
date)](https://img.shields.io/github/v/release/nmfs-fish-tools/pkgdownTemplate)](https://github.com/nmfs-fish-tools/pkgdownTemplate/releases)
[![R-CMD-check](https://github.com/nmfs-fish-tools/pkgdownTemplate/workflows/R-CMD-check/badge.svg)](https://github.com/nmfs-fish-tools/pkgdownTemplate/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


## Previous Training Agendas
[NSAW 2023 - Beginner TMB](https://nsawtraining.github.io/TMBtraining/articles/00-beginner-tmb-agenda.html)<br>
[NSAW 2023 - Intermediate TMB](https://nsawtraining.github.io/TMBtraining/articles/01-intermediate-tmb-agenda.html)


## Training Materials are found under the `docs` directory
| Folder | Description |
| -------|-------------|                                      
|[TMB Resources](https://nsawtraining.github.io/TMBtraining/articles/000_Getting_Started_Contents.html) | Links with helpful tips and resources on getting started with TMB |
|[Beginner TMB](https://nsawtraining.github.io/TMBtraining/articles/001_Beginner_TMB_Contents.html)|Materials for Beginners|
|[Intermediate TMB](https://nsawtraining.github.io/TMBtraining/articles/002_Intermediate_TMB_Contents.html)|Materials for Intermediate to Advanced users|

## Download materials

You can install materials from [GitHub](https://github.com/NSAWTraining/TMBtraining) with:

``` r
# install.packages("devtools")
devtools::install_github("NSAWTraining/TMBtraining")
```


<!-- Do not edit below. This adds the Disclaimer and NMFS footer. -->

****

```{r add-disclaimer, echo=FALSE, results='asis'}
url <- "https://raw.githubusercontent.com/nmfs-fish-tools/Resources/master/Disclaimer.md"
childtext <- readLines(url)
cat(childtext, sep="\n")
```

****

```{r footer, echo=FALSE, results='asis'}
url <- "https://raw.githubusercontent.com/nmfs-fish-tools/Resources/master/footer.md" 
childtext <- readLines(url)
cat(childtext, sep="\n")
```